\documentclass[b5paper]{book}
\input{common}
\usepackage{subfiles}
%\usepackage{layouts} %for printing the dimensions
\externaldocument{a}
\externaldocument{b}

\makeindex

\begin{document}

\newcommand{\titelpagina}[8]{%
\vspace*{4em}
\begin{center}
{\huge\sffamily\color{darkblue}The Category of Von Neumann Algebras}
\end{center}
\vspace{2em}
\begin{center}
{\par\noindent\sffamily\color{darkblue}\large\textbf{#1}}
\end{center}
\begin{center}
#2
\end{center}
\vspace{1em}
\begin{center}
#3
\end{center}
\begin{center}
#4
\end{center}
\begin{center}
#5
\end{center}
\begin{center}
#6
\end{center}
\vspace{10em}
\begin{center}%
#7
\end{center}%
\begin{center}%
\large\color{darkblue}Abraham Anton \textsc{Westerbaan}
\end{center}%
\begin{center}%
#8
\end{center}
\newpage
}

\newcommand{\achterkanttitelpagina}[3]{
\vspace*{8em}
\begin{center}
\textbf{\large\sffamily\color{darkblue}#1:}
\end{center}
\begin{center}
Prof.~dr.~B.P.F.~\textsc{Jacobs}
\end{center}
\vspace{5em}
\begin{center}
\textbf{\large\sffamily\color{darkblue}#2:}
\end{center}
\begin{center}
Prof.~dr.~J.D.M.~\textsc{Maassen}\\
\vspace{1em}
    Prof.~dr.~P.~\textsc{Panangaden} \\ {\footnotesize(McGill University, Canada)} \\
\vspace{1em}
    Prof.~dr.~P.~\textsc{Selinger} \\ {\footnotesize(Dalhousie University, Canada)} \\
\vspace{1em}
    Dr.~C.J.M.~\textsc{Heunen} \\ {\footnotesize(University of Edinburgh, #3)} \\
\vspace{1em}
Dr.~A.R.~\textsc{Kissinger}
\end{center}
\newpage
}

\newcommand{\whiteout}[1]{{\color{white}#1}}

\titelpagina{\whiteout{P}$\,$}{$\,$\whiteout{ter}\\
\whiteout{aan}$\,$\\
\whiteout{op}$\,$\\
\whiteout{volgens}$\,$\\
\whiteout{in}$\,$}{\whiteout{op}$\,$}{\whiteout{dinsdag}$\,$}{\whiteout{om}$\,$}{\whiteout{10.30}$\,$}{\whiteout{door}}{$\,$\whiteout{geboren}$\,$\\
\whiteout{te}$\,$}

\colofon{A.A.~Westerbaan}{978-94-6332-484-7}{1804.02203}{2066/201611}

\titelpagina{Proefschrift}{ter verkrijging van de graad van doctor\\
aan de Radboud Universiteit Nijmegen\\
op gezag van de rector magnificus prof. dr. J.H.J.M. \textsc{van Krieken},\\
volgens besluit van het college van decanen \\
in het openbaar te verdedigen}{op}{\textbf{dinsdag 14 mei 2019}}{om}{\textbf{10.30 uur precies}}{door}{geboren op 30 augustus 1988\\
te Nijmegen}

\achterkanttitelpagina{Promotor}{Manuscriptcommissie}{Verenigd Koninkrijk}

\titelpagina{Doctoral Thesis}{to obtain the degree of doctor \\
from  Radboud University Nijmegen \\
on the authority of the Rector Magnificus prof. dr. J.H.J.M. \textsc{van Krieken},\\
according to the decision of the Council of Deans \\
to be defended in public}{on}{\textbf{Tuesday, May 14, 2019}}{at}{\textbf{10.30 hours}}{by}{born on August 30, 1988\\
in Nijmegen (the Netherlands)}

\achterkanttitelpagina{Supervisor}{Doctoral Thesis Committee}{United Kingdom}

\vspace*{.1em}

\newpage
\makeatletter\@starttoc{parsectoc}\makeatother
\chapter{Introduction}

% Code to print the dimensions of the paper:
%\newpage
%\currentpage
%\pagedesign
%\currentpage
%\pagedesign

\begin{parsec}{10}
\begin{point}{10}
What does this Ph.D.~thesis offer?
Proof, perhaps,
to the doctoral thesis committee
of passable academic work;
an advertisement, as it may be,
of my school's perspective
to colleagues;
a display, even,
of intellectual achievement
to friends and family.
But I believe such narrow and selfish goals \emph{alone}
barely serve to keep a writer's spirits 
energised---and are definitely detrimental to that of the readers.
That is why I have foolhardily
challenged
myself
not just 
to drily list contributions,
but to write this thesis 
as the introduction,
that I would have liked to read
when I started
research for this thesis
back in May~2014.

The topic is von Neumann algebras,
the category they form,
and how they may be used
to model aspects of quantum computation.
Let us just say for now that a von Neumann algebra
is a special type of complex vector
space endowed with
a multiplication operation among some other additional structure.
An important example is the complex vector space~$M_2$
of~$2\times 2$ complex matrices,
because it models (the predicates on) a qubit;
but all~$N\times N$-complex matrices form a von Neumann algebra~$M_N$ as well.
Using von Neumann algebras
(and their little cousins, $C^*$-algebras) 
to describe quantum data types 
seems to be quite a recent idea
(see e.g.~\cite{jacobs2013block,rennela2015operator,furber2013kleisli}, 
	and~\cite{cho2016semantics} for an overview)
and has two distinct features.
Firstly, classical data types
are neatly incorporated:
$\C^2 \equiv \C\oplus \C$
models a bit,
and the direct sum $M_2\oplus M_3$
models the union type of a qubit and a qutrit.
Secondly,
von Neumann algebras
allow for infinite data types as well
	such as~$\scrB(\ell^2(\Z))$,
which represents a ``quantum integer.''\footnote{Though
	other methods of modelling infinite dimensional
	quantum computing have been proposed as well
	e.g.~using non-standard analysis\cite{Gogioso2017},
	pre-sheaves\cite{malherbe2013categorical},
	the geometry of interaction\cite{hasuo2017semantics},
	and quantitative semantics\cite{pagani2014applying}.}
It should be said that this last feature
is both a boon and a bane:
it brings with it all the inherent
intricacies of dealing with infinite dimensions;
and it is no wonder that
most authors choose 
to restrict themselves
to finite dimensions,
especially since
this seems to be enough to describe quantum algorithms,
see e.g.~\cite{nielsen2002quantum}.
\end{point}
\begin{point}{20}%
In this thesis, however,
we do face infinite dimensions,
because the two main results demand it:
\begin{enumerate}
\item
For the first result,
that von Neumann algebras
form  a model of Selinger and Valiron's quantum lambda calculus,
		as Cho and I explained in~\cite{model}
and for which I'll provide the foundation here,
we need to interpret function types,
some of which are essentially infinite dimensional.
\item
The second result,
an axiomatisation
of the map $a\mapsto \sqrt{p}a\sqrt{p}\colon \scrA\to\scrA$
representing measurement
of an element $p\in[0,1]_\scrA$
of a von Neumann algebra~$\scrA$
was tailored by B.E.~Westerbaan (my twin brother) and myself to work for
both finite and infinite dimensional~$\scrA$.
\end{enumerate}
These results
are part of a line of research that
tries to find patterns
in the category of von Neumann algebras,
that may also be cut from other categories
modelling computation---ideally in order to arrive at categorical axioms
for (probabilistic) computation in general.
When I joined the fray
the notion of \emph{effectus}\cite{newdirections} had already
been established by Jacobs,
and the two results
above offer potential additional axioms.
The work in this area has largely been a collaborative effort,
primarily between Jacobs, Cho, my twin brother, and myself,
and many of their insights have ended up in this thesis.

Of this I'd say no more
than that my work appears conversely, and proportionally,
in their writings too, except that the close cooperation
with my brother begs further explanation.
Our efforts on certain topics have been like interleaving 
of the pages of two phone books:
separating them  would be nigh impossible,
especially the work on the axiomatisation
of~$a\mapsto \sqrt{p}a\sqrt{p}$ and Paschke dilations.
So that's why we decided to write our theses
as two volumes of the same work;
preliminaries on von Neumann algebras,
and the axiomatisation of $a\mapsto \sqrt{p}a\sqrt{p}$
appear in this thesis,
while the work on dilations,
and effectus theory appear in my brother's thesis,
\cite{bas}.
\end{point}
\begin{point}{30}
The two results mentioned above only
make up 
about a third of this thesis;
the rest of it is devoted to 
the introduction to the theory of von Neumann algebras
needed to understand these results.
My aim is that anyone 
with, say, a bachelor's degree in mathematics
(more specifically, basic knowledge of linear algebra,
analysis\cite{rudin1964principles}, 
topology\cite{willard}
and set theory\cite{devlin2012joy})
should at least be able to follow the lines of reasoning
with only minimal recourse to external sources.
But I hope that they will gain some deeper understanding
of the material as well.
To this end, and because I wanted to gain some of this insight
for myself too,
I've not just mixed
and matched
results from the literature,
but I tailored a thorough treatise
of everything that's needed,
including proofs (except~\sref{intersection-tensor-proof}).
Whenever possible,
I've taken shortcuts
(e.g.~avoiding for example
the theory of Banach algebras
and locally convex spaces entirely)
to  prevent the mental tax
the added concepts
(and pages) would have brought.
For the same reasons
I've refrained from putting
everything in its proper abstract (and categorical\cite{maclane}) context
trusting that it'll shine through of its own accord.
I've however not been able to restrain
myself in making perhaps frivolous variations on the existing
theory whenever not strictly necessary,
taking for example Kadison's characterisation\cite{kadison1956}
of von Neumann algebras
as my definition,
and developing the elementary theory for it;
in my defence I'll just say this adds to
the original element that is expected of a thesis.
\end{point}
\begin{point}{40}{Advertisements}%
Due to space--time constraints
this thesis is based only on a selection
\cite{model,cho2015quotient,cho2016duplicable,qpakm,westerbaan2016universal}
of the works
I produced under supervision of Jacobs,
and while~\cite{wwpaschke,effintro,statesofconvexsets}
are incorporated in my brother's thesis,
this means~\cite{jacobs2015effect,jacobs2017distances} 
are unfortunately left out.
If you like this thesis,
then
you might also want to take a look
at these~\cite{rennela2017infinite,
rennela2015complete,
furber2013kleisli,
kornell2012,
heunen2015domains,
Maassen2010} recent works on von Neumann algebras,
and $C^*$-algebras.
If you're curious
about effectus theory
and related matters,
please have a look at~\cite{jacobs2017quantum,
cho2017disintegration,
jacobs2016hyper,
jacobs2017channel,
jacobs2017formal,
cho2017efprob,
jacobs2017probability,
jacobs2017recipe,
jacobs2016effectuses,
jacobs2016affine,
jacobs2016relating,
effintro,
statesofconvexsets,
cho2015quotient,
jacobs2017distances,
jacobs2015effect,
jacobs2016expectation,
jacobs2016predicate,
newdirections}.
But if you'd like more pictures instead,
I'd suggest~\cite{coecke2017picturing}.
\end{point}
\begin{point}{50}[on-writing-style]{Writing style}
I've replaced page numbers by
paragraph numbers
such as~\sref{on-writing-style}
for this paragraph.
The numbers after~\sref{final-bram} refer to paragraphs
	in my twin brother's thesis\cite{bas}.
\Define{Definitions}\index{definitions} are set like that
(i.e.~in blue),
and can be found in the index.
Proofs of certain facts
that are easily obtained on the back of an envelope,
and would clutter this manuscript,
have been left out.
Instead these facts have been phrased as exercises
as a challenge to the reader.
\end{point}
\begin{point}{51}
    The symbol ``\Define{$:=$}''\index{(((defequal@$:=$, is defined to be}
should be interpreted as ``is defined to be'',
    while ``\Define{$\equiv$}''\index{(((equiv@$\equiv$, being of the form}
should be read as ``being of the form''.
Sometimes ``$\equiv$'' is
used to define something on its right-hand side,
as in ``let $A\equiv\left(
\begin{smallmatrix} a&b\\b^* &c\end{smallmatrix}
\right)$ be a self-adjoint matrix.''
Other times~``$\equiv$''
indicates a simple rewrite step, as in
``since~$a=2$, we have $a+2=2+2\equiv 4$,''
where it's not suggested $a=2$ implies $2+2=4$.
\end{point}
\begin{point}{60}{Acknowledgements}
The work in this thesis specifically
has benefited greatly from
discussions
with John van de Wetering,
Robert Furber,
Kenta Cho,
and Bas Westerbaan,
but I've also had the pleasure
of discussing a variety
of other topics 
with 
Aleks Kissinger,
Andrew Polonsky,
Bert Lindenhovius,
Frank Roumen, 
Hans Maassen,
Henk Barendregt,
Joshua Moerman,
Martti Karvonen,
Robbert Krebbers,
Robin Adams, 
Robin Kaarsgaard,
Sam Staton, 
Sander Uijlen,
Sebastiaan Joosten,
and many others.
I'm especially honoured to have been received
in Edinburgh by Chris Heunen 
and in Oberwolfach
by Jianchao Wu.
I'm very grateful to
Arnoud van Rooij,
Bas Westerbaan
and
John van de Wetering
for proofreading large parts of
this manuscript,
without whose efforts
even more shameful errors would have remained.
I'm very grateful too for the manuscript committee's members' 
various suggestions and comments, and hope the improvements I made to this text
do them justice.
I should of course not forget to mention
the contribution
of friends (both close and distant),
family,
and colleagues---too numerous to name---of keeping me sane
these past years.

This is the second dissertation topic
I've worked on;
my first attempt
under different supervision
was unfortunately cut short after $1\sfrac{1}{2}$ years.
When Bart Jacobs graciously offered
me a second chance,
I initially had my reservations,
but accepted on account of the challenging topic.
Little did I know 
that behind the ambition and suit
	one finds a man
of singular moral fibre,
embodying
what was said
	about von Neumann himself:
	``[he] had to understand and accept much that most 
of us do not want to accept and do not even wish to understand.''%
\footnote{An excerpt from Eugene P.~Wigner's writings,
see page~130 of~\cite{wigner2013collected}.}
\end{point}
\begin{point}{70}{Funding} was received from the 
European Research Council under grant agreement \textnumero~320571.
\end{point}
\end{parsec}

%\subfile{apropos}
\subfile{cstar}
\subfile{vn}
\subfile{proc}
\begin{parsec}{1330}%
\begin{point}{10}[conclusion]{Conclusion}
Here ends this thesis,
but not the entire story.
There's much more to be said
about self-dual Hilbert $\scrA$-modules,
about dilations and their relation to purity,
and about the abstract theory of corners, filters,
and $\diamond$-positivity.
You'll see all this,
and more,
in the sequel,
``Dagger and dilations in the category of von Neumann algebras''\cite{bas},
brought to you by my twin brother.
\end{point}
\end{parsec}
\begin{parsec}{1340}[final-bram]
\begin{point}{10}
(Paragraphs numbered~\sref{final-bram}
and up can be found in~\cite{bas}.)
\end{point}
\end{parsec}


\backmatter

\fancyfoot[CE]{}
\fancyfoot[CO]{}
\fancypagestyle{plain}{
    \fancyfoot[CE]{}
    \fancyfoot[CO]{}
}

\printindex

\begingroup
\renewcommand\chapter[2]{\backmattertitle{Bibliography}}
\bibliography{main}{}
\endgroup

\bibliographystyle{plain}

\oldchapter{Lekensamenvatting}

Wat is wiskunde? Volgens mij is ze niets anders dan een studie van de patronen
in de wereld om ons heen in de meest \emph{strenge} en 
\emph{formele} zin.
De wiskundige
maakt zich  bewust, bijna komisch, blind 
voor zaken die een ander z\'o in het oog zouden springen.
``Ja.'' antwoorden op de vraag ``Wil je koffie of thee?'' is voor
een wiskundige volstrekt acceptabel.
Maar door deze blindheid kan de wiskundige wel
met een onge\"evenaarde nauwkeurigheid,
zekerheid en tijdloosheid
uitdrukking geven aan afzonderlijke  aspecten van de werkelijkheid.

Slechts een zeer beperkt aantal fenomenen
is vatbaar voor wiskundige \hyphenation{ana-lyse}analyse.
Soms begrijpen we een fenomeen er nog niet goed genoeg voor.
Soms is de benodigde wiskunde er nog niet voor ontwikkeld.
Maar in de meeste gevallen is het zo dat
het mes van de formele wiskunde simpelweg te scherp snijdt. 
Een wiskundige zou je niet kunnen vertellen
of een schip hetzelfde blijft als je alle planken vervangt,\footnote{Het schip
van Theseus, zie~\S23 van~\cite{theseus}.}
wat \emph{twee} is w\'el.
Het is niet zozeer dat hij of zij zich niet bezig zou willen houden
met dagelijkse begrippen, maar eerder dat deze begrippen daartegen niet
bestand zijn.
Daarom gaat de wiskunde over abstracte en ge\"idealiseerde 
objecten zoals getallen, vierkanten, cirkels, distributies,
tensoren en vari\"eteiten,
in plaats van wereldse tegenhangers zoals
respectievelijk
logistiek, akkers, hemellichamen, toekomstverwachtingen,
elektrische velden en sterrenstelsels.

Het is wonderbaarlijk hoe behulpzaam
wiskunde kan zijn ondanks en juist dankzij haar
beperkingen.
Niemand---ook geen wiskundige---kan zich het stilleven 
op de omslag van dit boekje vanuit een andere hoek nauwkeurig voorstellen.
Maar als ik vertel dat het een plaatje is van
een reflecterende bol in~$\R^{12}$
(tussen twee afgeknotte hypervlakken, met een deels reflecterend 
schaakbordpatroon) en wanneer ik 
de relevante co\"ordinaten en afmetingen geef, dan
kan ieder die \hyphenation{be-kend}bekend is met `inproducten' (vgl.~\sref{hilb-def})
een computerprogramma schrijven dat
dit tafereel vanuit een andere hoek toont.
Dankzij de wiskunde
is bovenmenselijk inzicht in de twaalfde dimensie niet nodig!

Dat sommige wiskundige begrippen zeer bruikbaar zijn,
wil niet zeggen dat \emph{elk} wiskundig begrip dit is.
Integendeel zelfs:
van alle denkbare figuren op papier
grijpt men ongetwijfeld het vaakst terug
naar de cirkel en de lijn.
Bepaalde constanten ($2\pi$, $e$, $\varphi$, $\sqrt{2}$, \ldots) 
verschijnen ook vaker in formules dan andere.
Natuurlijk kan een nuttig wiskundig begrip 
ook slecht gebruikt worden.
Men kan bijvoorbeeld 
de baan van een planeet met een groot
aantal cirkels (Ptolemaeus' `epicykels') 
omschrijven in plaats van met \'e\'en ellips.
Bij een dergelijk waardeoordeel als `goed' of `slecht' gebruik
van een begrip is er zeker sprake van persoonlijke
voorkeur, conventie en willekeur. Waarom gebruiken
we niet $2\pi=6,\!2831\dotsc$
als constante in plaats van
$\pi=3,\!1415\dotsc$? Volgens mij zou
dat veel formules eleganter maken;
zie~\cite{palais2001pi}.
Dat een wiskundig begrip (zoals epicykels)
in ongebruik geraakt is, wil bovendien niet zeggen
dat het geen \hyphenation{stijl-vol}stijlvol herintreden kan maken:
de befaamde Fourier-reeks (die in feite bestaat uit oneindig veel epicykels)
geeft de ellips weer 
het nakijken!\footnote{Of toch niet,  \cite{mayer2010drawing}?}

Dit proefschrift gaat over zo'n bruikbaar, abstract en 
ge\"idealiseerd wiskundig begrip:
de \emph{von Neumann-algebra},
bedacht door en vernoemd naar het Hongaarse genie
en de alleskunner
John von Neumann (geboren als Neumann J\'anos Lajos).
Aan hem hebben we niet alleen
de moderne computer-architectuur te danken,
maar bijvoorbeeld ook de springstoflenzen die in kernwapens gebruikt worden,
de afschrikwekkende strategie van \emph{mutual assured destruction}
en de numerieke weersvoorspelling.

Het is niet eenvoudig om uit te leggen
 wat von Neumann-algebra's precies zijn (\sref{vna}),
 maar 
ik kan wel een indruk geven waarvoor ik ze gebruik.
Uitgangspunt van dit proefschrift is dat 
de elementaire
systemen
die je tegenkomt bij het ontwerpen van een algoritme
voor een kwantumcomputer omschreven
kunnen worden door von Neumann-algebra's.
Zo wordt de klassieke bit (die~$0$ of~$1$ kan zijn)
voorgesteld door de von Neumann-algebra ``$\C^2$''
en wordt de kwantumbit (die een complexe combinatie van~$0$ of~$1$ is totdat
je haar meet)
voorgesteld door de von Neumann-algebra~``$M_2$''.
Dit geeft de hoop dat ingewikkeldere samengestelde systemen
ook een bijpassende von Neumann-algebra hebben.
Dat hangt natuurlijk af van de wijze van samenstelling:
het systeem dat bestaat uit twee kwantumbits en \'e\'en bit wordt voorgesteld
met de von Neumann-algebra $M_2\otimes M_2 \otimes \C^2$,
terwijl het systeem dat een kwantumbit of een klassieke bit bevat 
omschreven wordt met de von Neumann-algebra $M_2 \oplus \C^2$.
De bewerking ``$\otimes$'' op von Neumann-algebra's
is de interpretatie voor de  ``en''-samenstelling van systemen, 
terwijl ``$\oplus$''  invulling geeft aan ``of''-samenstelling van systemen.
Een veel complexere samenstelling van twee systemen~$A$ en~$B$
bestaat uit alle `processen' van~$A$ naar~$B$.
Meting is bijvoorbeeld zo'n proces van een kwantumbit naar een klassieke bit.
Zulke processen worden in de wereld van von Neumann-algebra's
voorgesteld door zogenaamde ``ncpsu-afbeeldingen'' (\sref{maps})
tussen von Neumann-algebra's.

E\'en van de hoofdresultaten van dit proefschrift
is een interpretatie voor deze `processen'-samenstelling,
de bewerking~$\limp$.
Door bepaalde formele kaders
was het al van tevoren duidelijk dat er hoogstens \'e\'en
interpretatie mogelijk zou zijn;
de vraag was alleen: welke?  Sterker nog: is er \"uberhaupt eentje?
Dit is vergelijkbaar met de vraag wat de kleinste\footnote{Kleinste
in de woordenboekordening.}
grammaticaal
correcte tekst is waarin
alle Nederlandse woorden voorkomen.
Om dit probleem op te lossen kun je niet zomaar alle woorden
op een rij zetten---de onderlinge samenhang moet immers kloppen.
De crux was voor mij om niet te willen proberen om de bewerking~$\limp$ 
direct te omschrijven,
maar om het bestaan ervan indirect aan te tonen,
zoals je ook kunt laten zien dat de  bovengenoemde Nederlandse tekst
uit het voorbeeld bestaat,
zonder te weten hoe hij precies is samengesteld.
(Voor een beperkte klasse van
von Neumann-algebra's, de `hereditair atomische', bleek een
directe omschrijving trouwens wel mogelijk, \sref{AstarhaB-concrete}.)

Het tweede hoofdresultaat van dit proefschrift
is de vondst van een abstracte omschrijving
van de processen
die bij een meting horen (\sref{sqrt-axiom}).
Het \hyphenation{ken-mer-ken-de}kenmerkende aspect van deze omschrijving is dat
het alleen gebruik maakt van zogenaamd `categorisch' jargon.
Dat dit mogelijk is toont aan
dat we in de categorie van von Neumann-algebra's
niet op een te abstract niveau werken:
we kunnen het nog steeds over meting hebben.
Dankzij de  categorische omschrijving  wordt het bovendien
mogelijk `meting' in andere contexten  te interpreteren.
Bij de zoektocht naar deze categorische omschrijving
was lange tijd het probleem om
een manier te vinden
om onderscheid te maken tussen de processen\footnote{Hier is~$u\neq 1$
een  unitaire
en 
$p$ een positief element van een von Neumann-algebra~$\scrA$
met $\ceil{p}=1$.}
\begin{equation*}
\sqrt{p}(\,\cdot\,)\sqrt{p}\qquad\text{en}\qquad
\sqrt{p}u^*(\,\cdot\,)u \sqrt{p}.
\end{equation*}
De linker hoort bij een meting, de rechter niet.
De oplossing was om een begrip uit de theorie 
van Hilbert-ruimten---geadjungeerdeerdheid---in een 
afgezwakte vorm over te nemen, namelijk $\diamond$-geadjungeerdheid 
(of \emph{contraposedness}, \sref{contraposed}).
Het linker proces blijkt een kwadraat te zijn van een aan zichzelf
$\diamond$-geadjungeerd proces; de rechter niet.
Beide hoofdresultaten staan in het laatste hoofdstuk.

De rest van dit proefschrift bestaat---enigszins ongebruikelijk---uit 
een grondige introductie
tot de benodigde, reeds bestaande, theorie
van $C^*$-algebra's en van von Neumann-algebra's.
Niet alleen bestond
een geschikte introductie nog niet, maar het leek me ook een goede kans 
om me verder te verdiepen in de theorie van von Neumann-algebra's.
In de eerste hoofdstukken ontwikkel ik de gehele benodigde theorie, inclusief 
bewijzen.
Het is mijn bedoeling dat eenieder die een bachelorgraad in de wiskunde heeft 
behaald deze tekst zou moeten kunnen begrijpen.
Op een groot aantal plekken wijk ik af van 
het begane pad:
soms om de tekst kort te houden, (zo ontwijk ik de
theorie van Banach algebra's volkomen),
maar meestal om te experimenteren met variaties.
Zo gebruik ik Kadison's omschrijving
van von Neumann-algebra's om
de theorie op te bouwen (wat niet eerder gedaan is.)
Zo houdt men goede wiskundige begrippen levend.

\oldchapter{About the Author}
Bram Westerbaan, born August 30, 1988,
enrolled as a physics and astronomy student at 
the Radboud University in 2006.
He obtained a bachelor's degree in mathematics in 2012 (cum laude)
with a thesis
on the simply typed $\lambda$-calculus
supervised by prof.~dr.~H.P.~Barendregt
(resulting in a publication~\cite{brambachelor}.)
The same year,
he obtained a master's degree in mathematics (summa cum laude)
with a thesis\cite{brammaster} on measure and integral
under supervision of prof.~dr.~A.C.M.~van Rooij.
His doctoral studies started in 2013
on the topic of (co)algebra guided by dr.~A.~Silva,
and shifted 
to the material presented in this thesis
in~2014
under the auspices of prof.~dr.~B.P.F.~Jacobs.
From 2018 onward Bram works as a postdoctoral researcher
at the digital security department on
an NWA project applying polymorphic pseudonymisation
to network traffic flow data.
\end{document}
