\documentclass[b5paper]{book}
\input{common}

\begin{document}
\begin{parsec}{10}%
A key to positivity is inversion --- both in  $C^*$-algebras,
and in life.
\end{parsec}
\begin{parsec}{20}
If we wish to purify our society from inexplicable intuitions,
then we should not base artificial intelligence on neural networks.
\end{parsec}
\begin{parsec}{30}%
The theory of von Neumann algebras
can be developed without mentioning
the spectral theorem.
\end{parsec}
\begin{parsec}{40}%
Breadth, not consistency, is a sign of truth;
quite the opposite.
\end{parsec}
\begin{parsec}{50}
Von Neumann algebras form a model of the quantum lambda calculus.
\end{parsec}
\begin{parsec}{60}
The sequential product can be axiomatised in the category of von Neumann
algebras.
\end{parsec}
\begin{parsec}{70}%
Von Neumann algebras
though 80 years old
need not be replaced.
\end{parsec}
\begin{parsec}{80}%
Forceful optimism
invites prevarication, silence,
and their pal, decay.
\end{parsec}
\begin{parsec}{90}%
Mediocrity is sticky.
\end{parsec}
\begin{parsec}{100}%
Taking offence where none was intended
is either just foolish\dots{} or cruel.
\end{parsec}
\begin{parsec}{110}%
A tragedy:
IQ is distributed normally,
but success is not.
\end{parsec}
\begin{parsec}{120}%
There's more to gardening
than adding and crossbreeding exciting new species.
We scholars should take note.
\end{parsec}
\begin{parsec}{130}%
A dissected animal is dead.
The same is true for concepts.
\end{parsec}
\begin{parsec}{140}%
Lies betray themselves by their
unnatural consistency.
\end{parsec}
\begin{parsec}{150}%
For all practical purposes
magic is a placeholder for better
articulated understanding.
\end{parsec}
\begin{parsec}{160}%
The fabric
of reality,
like the fabric of people,
bares no blame.
\end{parsec}
\begin{parsec}{170}%
In highly civilised society
there isn't quite enough serious input
for our hard-wired ``knight program'',
but the ``knave program'' comes along just fine.
\end{parsec}

\end{document}
